Query Optimization for Dynamic Imputation

### 2017-03-01 paper/relatedwork.tex
Change regression trees -> decision trees

Imputation of missing values is widely studied in the statistics and machine
learning communities. As highlighted in~\cite{gelman2006data}, missing data
can appear for a variety of reasons.
It can be missing uniformly at random or conditioned on
existing values (observed and missing). Methods in the statistical community
focus on correctly modeling relationships between attributes to account for different forms of missingness. For example, Burgette and Reiter~\cite{burgette2010multiple} discuss the usage of sequential regression trees
for imputing missing data.

~\cite{akande2015empirical} analyzes the performance of various
multiple imputation techniques on the American Community Survey dataset. 
The computational difficulties of imputing on large base
tables are well-known and can limit approaches. For example, Akande finds that
one approach (MI-GLM) is prohibitively expensive when attempting to impute on data that
includes variables with potentially large domains (ten categories in their case).
In contrast, \ProjectName{} allows users to specify a trade-off between
imputation quality and runtime performance, allowing users to perform queries
directly on the entire dataset. Furthermore,
the query planner's imputation is guided by the requirements of each specific
query's operators, rather than requiring broad assumptions about query
workloads.

\subsection{Missing Values and Databases}
There is a long history in the database community surrounding the
treatment of \nullv{}. As early as 1973,~\cite{codd1973understanding}
provided a treatment of the semantics of \nullv{}. Multiple
papers have described various (at times conflicting) treatments
of nulls~\cite{grant1977null}. \ProjectName's main design invariant---no relational operator
sees missing values for attributes it must operate on and users should never see
missing data---eliminates
the need to handle \nullv{} value semantics, while guaranteeing query evaluation soundness.

@@

In~\cite{wolf2007query}, queries over databases with missing values are
processed using a statistical approach, taking advantage of correlations between
attributes. The tuples that match the original query as well as a ranking of
tuples with incomplete data that may match are returned. Our work differs in
that we allow any well-formed statistical technique to be used for imputation
and focus on returning results to the analyst as if the database had been
complete.

Designers of data stream processing systems frequently confront missing values
and consider them carefully in query processing. Often, if sensor error is the
cause of missing values, values can be imputed with high confidence. In~\cite{fernandez2009inter}, feedback punctuation is used to dynamically
reconfigure the query plan for state-dependent
optimizations as data continue to arrive. One of the applications of this framework is to avoid
expensive imputations as real-time conditions change. 
\ProjectName{} focuses on handling
missing values, rather than providing a general framework for planning feedback. Furthermore,
it provides the ability
to trade imputation quality for runtime performance. 
%We prove our
%algorithm and cost model produces an optimal set of plans to choose from,
%given the constraints of the search space.

@@

\cite{duan2007processing} describes the Fa system and its declarative language for time-series forecasting. Their
system automatically searches the space of attribute combinations/transformations and statistical models
to produce forecasts within a given accuracy threshold. Accuracy estimates are determined using
standard techniques, such as k-fold cross-validation.

>>>>

Imputation of missing values is widely studied in the statistics and machine
learning communities. As highlighted in~\cite{gelman2006data}, missing data
can appear for a variety of reasons.
It can be missing uniformly at random or conditioned on
existing values (observed and missing). Methods in the statistical community
focus on correctly modeling relationships between attributes to account for different forms
of missingness. For example, \cite{burgette2010multiple} discuss the
usage of iterative decision trees for imputing missing data.

\cite{akande2015empirical} analyzes the performance of various
multiple imputation techniques on the American Community Survey dataset. 
The computational difficulties of imputing on large base
tables are well-known and can limit approaches. For example, \cite{akande2015empirical} finds that
one approach (MI-GLM) is prohibitively expensive when attempting to impute on data that
includes variables with potentially large domains (ten categories in their case).
In contrast, \ProjectName{} allows users to specify a trade-off between
imputation quality and runtime performance, allowing users to perform queries
directly on the entire dataset. Furthermore,
the query planner's imputation is guided by the requirements of each specific
query's operators, rather than requiring broad assumptions about query
workloads.

\subsection{Missing Values and Databases}
There is a long history in the database community surrounding the
treatment of \nullv{}. As early as 1973,~\cite{codd1973understanding}
provided a treatment of the semantics of \nullv{}. Multiple
papers have described various (at times conflicting) treatments
of \nullv{}s~\cite{grant1977null}. \ProjectName's main design invariant---no relational operator
sees missing values for attributes it must operate on nor do users see
missing data---eliminates
the need to handle \nullv{} value semantics, while guaranteeing query evaluation soundness.

@@

In~\cite{wolf2007query}, queries over databases with missing values are
processed directly using a statistical approach, taking advantage of correlations between
attributes. The tuples that match the original query as well as a ranking of
tuples with incomplete data that may match are returned. Our work differs in
that we allow any well-formed statistical technique to be used for imputation
and focus on returning results to the analyst as if the database had been
complete.

Designers of data stream processing systems frequently confront missing values
and consider them carefully in query processing. Often, if sensor error is the
cause of missing values, values can be imputed with high confidence. In~\cite{fernandez2009inter}, feedback punctuation is used to dynamically
reconfigure the query plan for state-dependent
optimizations as data continues to arrive. One of the applications of this framework is to avoid
expensive imputations as real-time conditions change. 
%We prove our
%algorithm and cost model produces an optimal set of plans to choose from,
%given the constraints of the search space.

@@

\cite{duan2007processing} describes the Fa system and its declarative language for time-series forecasting. Their
system automatically searches the space of attribute combinations/transformations and statistical models
to produce forecasts within a given accuracy threshold. Accuracy estimates are determined using
standard techniques, such as cross-validation.
